---
title: "covid-19"
author: "Jeonguk.lee"
date: "2025-07-29"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: cosmo
    highlight: tango
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 10,
  fig.height = 6
)
```

# COVID-19 Data Analysis

This document presents an analysis of COVID-19 data from Johns Hopkins University.
**Note**: This data was last updated on March 10, 2023.

## Library Loading

```{r libraries}
# Install and load required packages
packages <- c("readr", "dplyr", "ggplot2", "plotly", "DT", 
              "tidyr", "lubridate", "scales", "viridis", "zoo")

# Check and install missing packages
new_packages <- packages[!(packages %in% installed.packages()[,"Package"])]
if(length(new_packages)) install.packages(new_packages)

# Load libraries
library(readr)
library(dplyr)
library(ggplot2)
library(plotly)
library(DT)
library(tidyr)
library(lubridate)
library(scales)
library(viridis)
library(zoo)
```

## Data Loading

```{r data_loading}
# Correct raw GitHub URLs
urls <- c(
  us_confirmed = "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_US.csv",
  global_confirmed = "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv",
  us_deaths = "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_US.csv",
  global_deaths = "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv"
)

# Load data
cat("Loading data...\n")

global_confirmed <- read_csv(urls["global_confirmed"])
global_deaths <- read_csv(urls["global_deaths"])
us_confirmed <- read_csv(urls["us_confirmed"])
us_deaths <- read_csv(urls["us_deaths"])

cat("Data loading completed!\n")
```

## Data Structure Overview

```{r data_structure}
# Global confirmed cases data structure
cat("Global confirmed cases data structure:\n")
str(global_confirmed)

cat("\nFirst 5 columns and 3 rows:\n")
global_confirmed[1:3, 1:5] %>% 
  DT::datatable(options = list(scrollX = TRUE))
```

## Data Preprocessing

```{r data_preprocessing}
# Convert global confirmed cases data to long format
global_confirmed_long <- global_confirmed %>%
  pivot_longer(
    cols = -c(`Province/State`, `Country/Region`, Lat, Long),
    names_to = "date",
    values_to = "confirmed"
  ) %>%
  mutate(
    date = mdy(date),
    `Province/State` = replace_na(`Province/State`, "")
  ) %>%
  rename(
    province_state = `Province/State`,
    country_region = `Country/Region`
  )

# Convert global deaths data to long format
global_deaths_long <- global_deaths %>%
  pivot_longer(
    cols = -c(`Province/State`, `Country/Region`, Lat, Long),
    names_to = "date",
    values_to = "deaths"
  ) %>%
  mutate(
    date = mdy(date),
    `Province/State` = replace_na(`Province/State`, "")
  ) %>%
  rename(
    province_state = `Province/State`,
    country_region = `Country/Region`
  )

# Continental classification function
classify_continent <- function(country) {
  case_when(
    country %in% c("China", "Japan", "Korea, South", "Singapore", "Thailand", 
                   "Malaysia", "Philippines", "Vietnam", "Indonesia", "India", 
                   "Pakistan", "Bangladesh", "Sri Lanka", "Myanmar", "Cambodia",
                   "Laos", "Mongolia", "Nepal", "Bhutan", "Brunei", "Taiwan*",
                   "Afghanistan", "Kazakhstan", "Kyrgyzstan", "Tajikistan", 
                   "Turkmenistan", "Uzbekistan") ~ "Asia",
    
    country %in% c("Italy", "Spain", "Germany", "France", "United Kingdom",
                   "Netherlands", "Belgium", "Austria", "Switzerland", "Portugal",
                   "Sweden", "Norway", "Denmark", "Finland", "Poland", "Romania",
                   "Czech Republic", "Hungary", "Bulgaria", "Greece", "Croatia",
                   "Slovenia", "Slovakia", "Lithuania", "Latvia", "Estonia",
                   "Ireland", "Luxembourg", "Malta", "Cyprus", "Iceland",
                   "Russia", "Ukraine", "Belarus", "Moldova", "Serbia", "Bosnia and Herzegovina",
                   "Montenegro", "North Macedonia", "Albania", "Kosovo") ~ "Europe",
    
    country %in% c("US", "Canada", "Mexico", "Guatemala", "Belize", "El Salvador",
                   "Honduras", "Nicaragua", "Costa Rica", "Panama", "Cuba",
                   "Jamaica", "Haiti", "Dominican Republic", "Trinidad and Tobago",
                   "Barbados", "Bahamas", "Antigua and Barbuda", "Saint Lucia",
                   "Grenada", "Saint Vincent and the Grenadines", "Dominica",
                   "Saint Kitts and Nevis") ~ "North America",
    
    country %in% c("Brazil", "Argentina", "Chile", "Peru", "Colombia", "Venezuela",
                   "Ecuador", "Bolivia", "Uruguay", "Paraguay", "Guyana", 
                   "Suriname", "French Guiana") ~ "South America",
    
    country %in% c("Egypt", "South Africa", "Morocco", "Algeria", "Tunisia",
                   "Libya", "Sudan", "Ethiopia", "Kenya", "Uganda", "Tanzania",
                   "Rwanda", "Burundi", "Madagascar", "Mauritius", "Seychelles",
                   "Comoros", "Nigeria", "Ghana", "Senegal", "Mali", "Burkina Faso",
                   "Niger", "Chad", "Cameroon", "Central African Republic",
                   "Equatorial Guinea", "Gabon", "Congo (Brazzaville)", "Congo (Kinshasa)",
                   "Angola", "Zambia", "Zimbabwe", "Botswana", "Namibia", "Lesotho",
                   "Eswatini", "Malawi", "Mozambique", "Somalia", "Djibouti", "Eritrea") ~ "Africa",
    
    country %in% c("Australia", "New Zealand", "Papua New Guinea", "Fiji",
                   "Solomon Islands", "Vanuatu", "Samoa", "Tonga", "Kiribati",
                   "Tuvalu", "Nauru", "Palau", "Marshall Islands", "Micronesia") ~ "Oceania",
    
    TRUE ~ "Others"
  )
}

# Combine confirmed and deaths data with continental information
global_combined <- global_confirmed_long %>%
  left_join(
    global_deaths_long %>% select(province_state, country_region, date, deaths),
    by = c("province_state", "country_region", "date")
  ) %>%
  filter(!is.na(date)) %>%
  mutate(continent = classify_continent(country_region))

cat("Preprocessing completed! Total", nrow(global_combined), "records\n")
cat("Data period:", min(global_combined$date), "to", max(global_combined$date), "\n")
cat("Continental classifications:", unique(global_combined$continent), "\n")
```

## Global Overview

```{r global_summary}
# Daily global cumulative statistics
world_daily <- global_combined %>%
  group_by(date) %>%
  summarise(
    total_confirmed = sum(confirmed, na.rm = TRUE),
    total_deaths = sum(deaths, na.rm = TRUE),
    .groups = 'drop'
  ) %>%
  mutate(
    daily_new_cases = total_confirmed - lag(total_confirmed, default = 0),
    daily_new_deaths = total_deaths - lag(total_deaths, default = 0),
    case_fatality_rate = round(total_deaths / total_confirmed * 100, 2)
  )

# Latest statistics
latest_stats <- world_daily %>% 
  slice_tail(n = 1)

cat("=== Global Final Statistics (", as.character(latest_stats$date), ") ===\n")
cat("Total Confirmed Cases:", format(latest_stats$total_confirmed, big.mark = ","), "\n")
cat("Total Deaths:", format(latest_stats$total_deaths, big.mark = ","), "\n")
cat("Case Fatality Rate:", latest_stats$case_fatality_rate, "%\n")
```

## Visualizations

### 1. Global Cumulative Cases and Deaths Trends

```{r global_trends}
p1 <- world_daily %>%
  select(date, total_confirmed, total_deaths) %>%
  pivot_longer(cols = c(total_confirmed, total_deaths), 
               names_to = "type", values_to = "count") %>%
  mutate(
    type = case_when(
      type == "total_confirmed" ~ "Confirmed Cases",
      type == "total_deaths" ~ "Deaths"
    )
  ) %>%
  ggplot(aes(x = date, y = count, color = type)) +
  geom_line(size = 1) +
  scale_y_continuous(labels = comma_format()) +
  scale_color_manual(values = c("Confirmed Cases" = "#E31A1C", "Deaths" = "#1F78B4")) +
  labs(
    title = "Global COVID-19 Cumulative Cases and Deaths Trends",
    subtitle = paste("Data period:", min(world_daily$date), "to", max(world_daily$date)),
    x = "Date",
    y = "Cumulative Count",
    color = "Type"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 10),
    legend.position = "bottom"
  )

ggplotly(p1)
```

### 2. Daily New Cases Trends

```{r daily_new_cases}
p2 <- world_daily %>%
  filter(date >= as.Date("2020-03-01")) %>%  # Start from March
  ggplot(aes(x = date, y = daily_new_cases)) +
  geom_col(fill = "#FF6B6B", alpha = 0.7) +
  geom_smooth(method = "loess", se = FALSE, color = "#4ECDC4", size = 1) +
  scale_y_continuous(labels = comma_format()) +
  labs(
    title = "Global Daily New COVID-19 Cases Trends",
    subtitle = "Smooth curve represents trend line",
    x = "Date",
    y = "Daily New Cases"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold")
  )

ggplotly(p2)
```

### 3. Major Countries Comparison

```{r country_comparison}
# Top 10 countries by confirmed cases
top_countries <- global_combined %>%
  filter(date == max(date)) %>%
  group_by(country_region) %>%
  summarise(total_confirmed = sum(confirmed, na.rm = TRUE), .groups = 'drop') %>%
  arrange(desc(total_confirmed)) %>%
  slice_head(n = 10) %>%
  pull(country_region)

# Trends for top 10 countries
country_trends <- global_combined %>%
  filter(country_region %in% top_countries) %>%
  group_by(country_region, date) %>%
  summarise(
    confirmed = sum(confirmed, na.rm = TRUE),
    deaths = sum(deaths, na.rm = TRUE),
    .groups = 'drop'
  )

p3 <- country_trends %>%
  ggplot(aes(x = date, y = confirmed, color = country_region)) +
  geom_line(size = 1) +
  scale_y_log10(labels = comma_format()) +
  scale_color_viridis_d(name = "Country") +
  labs(
    title = "COVID-19 Cases Trends by Major Countries (Log Scale)",
    subtitle = "Top 10 countries by confirmed cases",
    x = "Date",
    y = "Cumulative Confirmed Cases (Log Scale)"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    legend.position = "bottom"
  ) +
  guides(color = guide_legend(nrow = 2))

ggplotly(p3)
```

### 4. Country-wise Final Statistics Table

```{r country_table}
final_country_stats <- global_combined %>%
  filter(date == max(date)) %>%
  group_by(country_region) %>%
  summarise(
    confirmed_cases = sum(confirmed, na.rm = TRUE),
    deaths = sum(deaths, na.rm = TRUE),
    .groups = 'drop'
  ) %>%
  mutate(
    fatality_rate = round(deaths / confirmed_cases * 100, 2)
  ) %>%
  arrange(desc(confirmed_cases)) %>%
  rename(country = country_region)

# Display top 20 countries
final_country_stats %>%
  slice_head(n = 20) %>%
  DT::datatable(
    caption = "COVID-19 Final Statistics by Country (Top 20)",
    options = list(
      pageLength = 10,
      scrollX = TRUE
    )
  ) %>%
  DT::formatCurrency(c("confirmed_cases", "deaths"), currency = "", interval = 3, mark = ",", digits = 0) %>%
  DT::formatRound("fatality_rate", digits = 2)
```

## South Korea Detailed Analysis

```{r korea_analysis}
# Extract South Korea data
korea_data <- global_combined %>%
  filter(country_region == "Korea, South") %>%
  group_by(date) %>%
  summarise(
    confirmed = sum(confirmed, na.rm = TRUE),
    deaths = sum(deaths, na.rm = TRUE),
    .groups = 'drop'
  ) %>%
  mutate(
    daily_new_cases = confirmed - lag(confirmed, default = 0),
    daily_new_deaths = deaths - lag(deaths, default = 0)
  )

# South Korea statistics
korea_latest <- korea_data %>% slice_tail(n = 1)

cat("=== South Korea Final Statistics ===\n")
cat("Total Confirmed Cases:", format(korea_latest$confirmed, big.mark = ","), "\n")
cat("Total Deaths:", format(korea_latest$deaths, big.mark = ","), "\n")
cat("Case Fatality Rate:", round(korea_latest$deaths / korea_latest$confirmed * 100, 2), "%\n")

# South Korea daily new cases chart
p4 <- korea_data %>%
  filter(date >= as.Date("2020-02-15")) %>%
  ggplot(aes(x = date, y = daily_new_cases)) +
  geom_col(fill = "#FF9999", alpha = 0.7) +
  geom_smooth(method = "loess", se = FALSE, color = "#0066CC", size = 1) +
  scale_y_continuous(labels = comma_format()) +
  labs(
    title = "South Korea Daily New COVID-19 Cases",
    subtitle = "From February 15, 2020",
    x = "Date",
    y = "Daily New Cases"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold")
  )

ggplotly(p4)
```

## Continental Comparison Analysis

```{r continent_analysis}
# Daily cumulative statistics by continent
continent_daily <- global_combined %>%
  group_by(continent, date) %>%
  summarise(
    confirmed = sum(confirmed, na.rm = TRUE),
    deaths = sum(deaths, na.rm = TRUE),
    .groups = 'drop'
  ) %>%
  filter(continent != "Others")  # Exclude others

# Final statistics by continent
continent_final <- continent_daily %>%
  group_by(continent) %>%
  slice_tail(n = 1) %>%
  arrange(desc(confirmed)) %>%
  mutate(
    fatality_rate = round(deaths / confirmed * 100, 2)
  )

cat("=== Final Statistics by Continent ===\n")
for(i in 1:nrow(continent_final)) {
  cat(continent_final$continent[i], ": Confirmed", 
      format(continent_final$confirmed[i], big.mark = ","), ", Deaths",
      format(continent_final$deaths[i], big.mark = ","), ", Fatality Rate",
      continent_final$fatality_rate[i], "%\n")
}
```

### Continental Confirmed Cases Trends

```{r continent_confirmed_trends}
p5 <- continent_daily %>%
  ggplot(aes(x = date, y = confirmed, color = continent)) +
  geom_line(size = 1.2) +
  scale_y_continuous(labels = comma_format()) +
  scale_color_viridis_d(name = "Continent") +
  labs(
    title = "COVID-19 Cumulative Cases Trends by Continent",
    subtitle = "January 2020 to March 2023",
    x = "Date",
    y = "Cumulative Confirmed Cases"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    legend.position = "bottom"
  )

ggplotly(p5)
```

### Continental Fatality Rate Comparison

```{r continent_fatality_rate}
p6 <- continent_final %>%
  ggplot(aes(x = reorder(continent, confirmed), y = fatality_rate, fill = continent)) +
  geom_col(alpha = 0.8) +
  geom_text(aes(label = paste0(fatality_rate, "%")), 
            hjust = -0.1, size = 3.5) +
  coord_flip() +
  scale_fill_viridis_d(guide = "none") +
  labs(
    title = "COVID-19 Fatality Rate Comparison by Continent",
    subtitle = "Based on final cumulative data",
    x = "Continent",
    y = "Fatality Rate (%)"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold")
  )

ggplotly(p6)
```

## Cases-Deaths Gap Analysis

```{r gap_analysis}
# Global case fatality rate changes over time
world_cfr_trends <- world_daily %>%
  filter(total_confirmed >= 1000) %>%  # Analyze from 1000+ cases
  mutate(
    cfr_7day = zoo::rollmean(case_fatality_rate, k=7, fill=NA, align="right"),
    confirmed_growth_rate = (total_confirmed / lag(total_confirmed, 7) - 1) * 100,
    deaths_growth_rate = (total_deaths / lag(total_deaths, 7) - 1) * 100
  )

# Case fatality rate trends
p7 <- world_cfr_trends %>%
  ggplot(aes(x = date)) +
  geom_line(aes(y = case_fatality_rate), color = "#FF6B6B", alpha = 0.3) +
  geom_line(aes(y = cfr_7day), color = "#E31A1C", size = 1.2) +
  scale_y_continuous(
    labels = function(x) paste0(x, "%"),
    limits = c(0, max(world_cfr_trends$case_fatality_rate, na.rm = TRUE) * 1.1)
  ) +
  labs(
    title = "Global COVID-19 Case Fatality Rate Trends",
    subtitle = "Red line: 7-day moving average, Light line: Daily fatality rate",
    x = "Date",
    y = "Case Fatality Rate (%)",
    caption = "Analysis starts from 1,000+ confirmed cases"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold")
  )

ggplotly(p7)
```

### Cases vs Deaths Ratio Over Time

```{r ratio_analysis}
# Comparison of relative growth rates for cases and deaths
p8 <- world_cfr_trends %>%
  filter(!is.na(confirmed_growth_rate), !is.na(deaths_growth_rate)) %>%
  select(date, confirmed_growth_rate, deaths_growth_rate) %>%
  pivot_longer(cols = c(confirmed_growth_rate, deaths_growth_rate),
               names_to = "type", values_to = "growth_rate") %>%
  mutate(
    type = case_when(
      type == "confirmed_growth_rate" ~ "Cases Growth Rate",
      type == "deaths_growth_rate" ~ "Deaths Growth Rate"
    )
  ) %>%
  ggplot(aes(x = date, y = growth_rate, color = type)) +
  geom_line(size = 1) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
  scale_color_manual(values = c("Cases Growth Rate" = "#4ECDC4", "Deaths Growth Rate" = "#FF9999")) +
  labs(
    title = "Cases vs Deaths Weekly Growth Rate Comparison",
    subtitle = "Positive: increase, Negative: decrease (7-day basis)",
    x = "Date",
    y = "Weekly Growth Rate (%)",
    color = "Type"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    legend.position = "bottom"
  )

ggplotly(p8)
```

### Cases-Deaths Gap Visualization

```{r gap_visualization}
# Gap between cumulative cases and deaths
p9 <- world_daily %>%
  mutate(gap = total_confirmed - total_deaths) %>%
  ggplot(aes(x = date)) +
  geom_area(aes(y = total_confirmed), fill = "#FFE5E5", alpha = 0.7) +
  geom_area(aes(y = total_deaths), fill = "#FF6B6B", alpha = 0.8) +
  geom_line(aes(y = gap), color = "#0066CC", size = 1.2) +
  scale_y_continuous(labels = comma_format()) +
  labs(
    title = "Global COVID-19 Cases-Deaths Gap Evolution",
    subtitle = "Red area: Cumulative deaths, Pink area: Cumulative cases, Blue line: Gap",
    x = "Date",
    y = "Cumulative Count",
    caption = "Widening gap indicates improving survival rates"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold")
  )

ggplotly(p9)
```

### Detailed Case Fatality Rate Analysis by Period

```{r detailed_cfr_analysis}
# Case fatality rate analysis by major periods
cfr_periods <- world_cfr_trends %>%
  mutate(
    period = case_when(
      date < as.Date("2020-06-01") ~ "Early Phase (Jan-May 2020)",
      date < as.Date("2020-12-01") ~ "First Wave (Jun-Nov 2020)",
      date < as.Date("2021-06-01") ~ "Vaccine Introduction (Dec 2020-May 2021)",
      date < as.Date("2021-12-01") ~ "Delta Variant (Jun-Nov 2021)",
      date < as.Date("2022-06-01") ~ "Omicron Variant (Dec 2021-May 2022)",
      TRUE ~ "Endemic Transition (Jun 2022 onwards)"
    )
  ) %>%
  filter(!is.na(cfr_7day))

# Average fatality rate by period
period_summary <- cfr_periods %>%
  group_by(period) %>%
  summarise(
    avg_cfr = mean(cfr_7day, na.rm = TRUE),
    min_cfr = min(cfr_7day, na.rm = TRUE),
    max_cfr = max(cfr_7day, na.rm = TRUE),
    .groups = 'drop'
  ) %>%
  arrange(avg_cfr)

# Box plot of fatality rates by period
p10 <- cfr_periods %>%
  ggplot(aes(x = reorder(period, -cfr_7day, FUN = median), y = cfr_7day, fill = period)) +
  geom_boxplot(alpha = 0.7) +
  geom_point(alpha = 0.3, position = position_jitter(width = 0.2)) +
  coord_flip() +
  scale_fill_viridis_d(guide = "none") +
  labs(
    title = "COVID-19 Case Fatality Rate Distribution by Period",
    subtitle = "Distribution of 7-day moving average fatality rates for each period",
    x = "Period",
    y = "Case Fatality Rate (%)"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold")
  )

ggplotly(p10)

# Period statistics table
period_summary %>%
  mutate(
    average_cfr = round(avg_cfr, 2),
    minimum_cfr = round(min_cfr, 2),
    maximum_cfr = round(max_cfr, 2)
  ) %>%
  select(Period = period, average_cfr, minimum_cfr, maximum_cfr) %>%
  DT::datatable(
    caption = "Case Fatality Rate Statistics by Period",
    options = list(
      pageLength = 10
    )
  )
```

## Key Findings Summary

```{r key_findings}
# Calculate key statistics
max_daily_cases <- max(world_daily$daily_new_cases, na.rm = TRUE)
max_daily_cases_date <- world_daily$date[which.max(world_daily$daily_new_cases)]

initial_cfr <- world_cfr_trends$case_fatality_rate[1]
final_cfr <- tail(world_cfr_trends$case_fatality_rate, 1)
cfr_improvement <- round((initial_cfr - final_cfr), 2)

# Continental ranking
continent_ranking <- continent_final %>%
  arrange(desc(confirmed)) %>%
  pull(continent)

cat("=== Key Findings ===\n\n")
cat("📊 **Global Peak Daily New Cases**\n")
cat("   ", format(max_daily_cases, big.mark = ","), "cases on", as.character(max_daily_cases_date), "\n\n")

cat("📈 **Case Fatality Rate Evolution**\n")
cat("   Initial CFR:", round(initial_cfr, 2), "% → Final CFR:", round(final_cfr, 2), "%\n")
cat("   Improvement:", cfr_improvement, "percentage points\n\n")

cat("🌍 **Continental Ranking by Cases**\n")
for(i in 1:length(continent_ranking)) {
  cases <- continent_final$confirmed[continent_final$continent == continent_ranking[i]]
  cat("   ", i, ".", continent_ranking[i], "-", format(cases, big.mark = ","), "cases\n")
}
```

### 💡 Analysis Interpretation

**1. Dramatic Improvement in Case Fatality Rate**
- From initial `r round(initial_cfr, 1)`% to final `r round(final_cfr, 1)`%, an improvement of `r cfr_improvement` percentage points
- This reflects the combined effect of medical experience accumulation, treatment development, and vaccine introduction

**2. Continuous Widening of Cases-Deaths Gap**
- Over time, while cases increased, the death rate increase was relatively slower
- The gap expanded rapidly especially after vaccination began in 2021

**3. Continental Characteristics Differences**
- Asia: Relatively low fatality rate compared to population → Effect of mask culture and early response
- Europe: High cases and fatality rates → Impact of delayed initial response and population density
- North America: High absolute numbers due to US influence

**4. Distinct Patterns by Period**
- **Early (2020)**: High fatality rate, healthcare system burden
- **Middle (2021)**: Rapid decrease in fatality rate due to vaccine introduction
- **Late (2022~)**: Many cases but relatively low fatality rate due to Omicron variant

---

*📈 This analysis provides an important record of how the COVID-19 pandemic evolved and how humanity adapted to it.*

## Conclusions

This analysis reveals several key insights about the COVID-19 pandemic:

### Global Overview
- **Total Scale**: `r format(latest_stats$total_confirmed, big.mark = ",")` confirmed cases and `r format(latest_stats$total_deaths, big.mark = ",")` deaths occurred
- **Final Case Fatality Rate**: Global average fatality rate was approximately `r latest_stats$case_fatality_rate`%

### Continental Characteristics
- **Asia**: High absolute case numbers but relatively low fatality rates
- **Europe**: High case numbers and relatively high fatality rates
- **North America**: High case numbers due to US influence
- **South America, Africa**: Recorded relatively high fatality rates

### Temporal Evolution Patterns
1. **Initial High Fatality**: Early pandemic (first half of 2020) had fatality rates above 5%
2. **Gradual Improvement**: Continuous decrease in fatality rates due to treatment development and healthcare system adaptation
3. **Vaccine Effect**: Rapid decrease in fatality rates after vaccine introduction in 2021
4. **Variant Impact**: Temporary increase during Delta variant period, sharp decrease during Omicron period

### Widening Cases-Deaths Gap
- **Continuous Gap Expansion**: Over time, the ratio of deaths to cases decreased
- **Medical Advancement**: Result of improved treatments, vaccination, and accumulated medical experience
- **Variant Characteristics**: Impact of variants with relatively lower fatality rates like Omicron

### South Korea's Characteristics
- Total of `r format(korea_latest$confirmed, big.mark = ",")` confirmed cases
- Demonstrated effective quarantine policies with relatively low fatality rates
- Proved initial response and healthcare system efficiency

### Data Limitations
- This data is only updated until March 10, 2023, requiring other sources for latest trends
- Potential data bias due to differences in testing policies and reporting standards by country

---

*Data Source: Johns Hopkins University Center for Systems Science and Engineering (JHU CSSE)*  
*Analysis Date: `r Sys.time()`*